---
title: "Homework 2"
author: "46-921, Fall 2018"
date: "Due Wednesday, September 12 at 1:00 PM"
output: pdf_document
---

\large

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

You should submit the Rmd file with your answers in the
appropriate spaces. Rename the file as `YOURANDREWID_HW2.Rmd`
and submit it via Canvas. Also submit the `.pdf` file that
is produced.

Of course, any code that is written should be tested. In each
case you are requested to include some __simple__ examples showing that
your code works. Think about how your example(s) can illustrate 
the range of possibilities that the code could face.

\vspace{.2in}

1. Suppose that `charvec` is a vector of type `character`.
Write a __single__ line of R code that returns a new, single character
string that consists of all of the elements of `charvec`
concatenated end to end. For example,
`charvec = c("a","b","c")`
should become `"abc"`.

```{r}
charvec1 = c("a","b","c")
charvec2 = c("a", " ", "x")
charvec3 = c("This", "is","a","Test")
paste(charvec1,sep='',collapse = '')
paste(charvec2,sep='',collapse = '')
paste(charvec3,sep='',collapse = '')
```
This paste can be used on string data after reading it from a csv file, to combine multiple columns into one. It can also be used for labeling plots made in R

2. Write an R function that takes a numeric vector as input
and returns the ``five-number summary`` as a __list__.
The five-number summary consists of the minimum, the
25th, 50th, and 75th percentile, along with the maximum.
Name the components of the list appropriately.

```{r}
five_num_summary = function(numvec)
{
   res = quantile(numvec,c(0.0, 0.25, 0.5, 0.75, 1.0),na.rm = T)
   return( as.list(res))
}
five_num_summary(c(1,2,3,4,5,NA,6,7,8,9,10))
five_num_summary(c(1.1,2.2,3.3,4.4,5.5,NA,6.6,NA,7.7))
```
This function can be called on a large dataset to quickly get the useful quantile information.


3. Write an R function that, when given a vector of strings,
returns the position(s) of the longest strings in the vector.

```{r}
longest_string_pos = function(stringvec)
{
   countcharvec = sapply(stringvec,nchar, USE.NAMES = F)
   pos = which(countcharvec ==  max(countcharvec))
   return(pos)
}
longest_string_pos(c("one", "two", "three", "five", "eleven"))
longest_string_pos(c("This", "is", "a", "test", "for", "multple", "longest", "string"))
longest_string_pos(c("Does", "this", "Work", "????"))
```

4. Create an __infix__ operator that returns whether or not a
number is a multiple of another. Call it %m%. In other words,
`10 %m% 5` should be `TRUE` while `10 %m% 3` should be `FALSE`.

```{r}
"%m%" = function(x,y)
{
  if(x%%y == 0 || y%%x == 0)
    return(TRUE)
  
  return(FALSE)
}

10%m%5
5%m%10
1%m%10        #Should be true always
1.23%m%3.69   #Works for fractions too!
1.24%m%5      #Should return false
```


5. Go to the following website:
[https://www.sec.gov/data/foiadocsfailsdatahtm](https://www.sec.gov/data/foiadocsfailsdatahtm)
and download the __July 2017, Second Half__ data set.
Read these data into R and create an appropriate data frame.
Be sure that each column is in an appropriate form (date, factor,
character, etc.)
When reading in the file, __do not make any changes to the data
file itself__. Resolve any issues using appropriate R commands.

```{r}
data = read.delim("cnsfails201707b.txt", header = TRUE, sep = "|")
data$SETTLEMENT.DATE = as.Date(as.character(data$SETTLEMENT.DATE),format = "%Y%m%d")
df = as.data.frame(data)
```

