---
title: "Homework 5"
author: "46-923, Fall 2018"
date: "Due Tuesday, December 4 at 2:45 PM"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

\large

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
You should complete this homework solely within R Markdown. You do not need to show
any derivations that were required to complete Question 1, just implement the analysis in R.
Question 1
Assume that X1,X2, . . . ,Xn are iid from the Poisson(lambda) distribution.
Assume you want to test H0 : lambda = 4.3 versus H1 : lambda < 4.3 using the Wald test. Calculate the
p-value when the data are those found via
x = read.table("http://www.stat.cmu.edu/~cschafer/MSCF/PoisHypTest.txt")
\vspace{.2in}

Answer: Under the H0, standard error of the mean of the poisson sample is sqrt(lambda_0/n). T-stat for the lambda_hat is (lambda_hat-lambda_0)/S.E, which is -0.38.p-value for the one sided test for this T-stat is 0.3515. p-value for this sample observation is very large and under standard confidence levels the null hypothesis cannot be rejected.

```{r}
x = read.table("http://www.stat.cmu.edu/~cschafer/MSCF/PoisHypTest.txt")
lambda_hat = mean(x$V1)
lambda_0   = 4.3
SE = sqrt(lambda_0/nrow(x))
Z = (lambda_hat-lambda_0)/SE
p_val = pnorm(Z)
```
Question 2

The Augmented Dickey-Fuller (ADF) Test is a test utilized in time series analysis in order to assess stationarity. There is a function adf.test() in the package tseries which implements this test. Take a look at help(adf.test) for some details.

a. If our objective is to see if there is strong evidence that a time series is stationary, how
should the argument alternative be set when using adf.test()?

Answer: 


b. Consider the following R commands. This will read in the stock data discussed in lecture,
and then run the ADF test on the first stock in the sample. Use this as a starting point
to write code to loop over all 1000 stocks and get the p-values for each test. How many
of the tests have p-value less than 0.05?
stocksample = read.table("stocksample.txt", header=T,
sep="\t", comment.char="")
adf.test(as.numeric(log(stocksample[1,5:34])), alternative="s")
c. Run the p-values found in part (b) through the Benjamini-Hochberg procedure described
in lecture. Are any of the series found to be stationary using this approach (again using
 = 0.05)? Comment on the reason(s) for any differences found.


```{r}
stocksample = read.table("stocksample.txt", header=T,
sep="\t", comment.char="")
adf.test(as.numeric(log(stocksample[1,5:34])), alternative="s")
```
