---
title: "Final Exam - Part One"
author: "46-921, Fall 2018"
date: "Friday, October 19"
output: pdf_document
---

\large

```{r setup, include=FALSE,message=FALSE}
#setwd("~/Teaching/MSCF/DScourses/FinalExam/Mini1/2018")
knitr::opts_chunk$set(echo = TRUE)
library(quantmod)
library(ggplot2)
```

#General Instructions

Please read the following carefully:

1. You have three hours total to work on the exam (both Parts One and Two), starting at
10:00 AM and ending at 1:00 PM. You will have to stop
working on the exam at 1:00 PM regardless of what time you
started working on the exam.

2. You are allowed to use any notes, calculators, web sites,
etc. that you wish, but you are not allowed to listen to any
audio, and you are __not allowed to communicate with anyone
else__, regardless of whether or not they are a student in the
course.

3. Remember that plots must be __readable__ and axes appropriately
labelled.

4. You are to turn in this .Rmd file and the corresponding
.pdf file. __Once you turn in Part One, then Part Two will become
available to you.__

5. Only your __first__ submission of Part One will be graded.

6. In all cases, please __be clear__ in your steps and your
explanations.

7. You can use any R packages and/or functions that you deem
helpful. We need to be able to reproduce your code, however.
We will install R packages, if needed.

8. Please name your submission `YOURANDREWID_PARTONE.Rmd`.

9. The number of points for
each question is shown in bold at the end of the question.
Everyone receives 20 points for free. Part One is worth
20 points. Part Two is worth 60 points.

\newpage


#The Data

We will use the "Blockholders" data set obtained from _Wharton Research Data Services_.
Their description is as follows:

"This dataset contains standardized data for blockholders of 1,913 companies. The data was cleaned from biases and mistakes usually observed in the standard source for this particular type of data. Blockholders' data is reported by firm for the period 1996-2001. The data cleaning procedure is explained in detail by Jennifer Dlugosz, Rudiger Fahlenbrach, Paul A. Gompers, and Andrew Metrick in their study 'Large Blocks of Stocks: Prevalance, Size, and Measurement'."

A _blockholder_ is a general term for a shareholder who
holds a large number of shares in that company, and hence
has a relatively large influence on company decisions.

The data file can be found on Canvas, in the "Final Exam"
folder under "Files." The name is `blockholders.csv`.

A list of variables is provided on the following page. 

\newpage

\#           Name         Description
-----------  ------       -----------
1	           `CompName`   Company Name
2	           `group2`     category of percentage blockholdings (0, 5-10, 10-15, 15-25, 25-50, >50)
3	           `SH_name`    Name of the blockholder as identified in the proxy statement
4	           `mtgdate`    Date of the annual proxy meeting
5	           `shrsrcd`    Date the Def14 A was filed with the SEC.
6	           `Ticker`     Ticker Symbol
7	           `Partial`    Corrected for partial overlap in raw data
8	           `Full`       Corrected for Full overlap situation in raw data
9	           `PrefFlag`   Corrected for Wrong attribution of preferred shares in raw data
10	         `irrcyear`   Year of the annual meeting
11	         `other`      Corrected for Wrong entry in raw data
12	         `sumblks`    Percentage held by all blockholders for that firm-year
13	         `numblks`    Number of all blockholders for that firm-year
14	         `sumaflin`   Percentage held by all affiliated blockholders
15	         `sumout`     Percentage held by all outside blockholders
16	         `sumesop`    Percentage held by all ESOP-related blockholders
17	         `sumdir`     Percentage held by all non-officer director blockholders
18	         `sumoff`     Percentage held by all officer blockholders
19	         `numaflin`   Number of affiliated blockholders
20	         `numout`     Number of outside blockholders
21	         `numesop`    Number of ESOP (Employee Share Ownership Plans) blockholders
22	         `numdir`     Number of non-officer director blockholders
23	         `numoff`     Number of officer blockholders
24	         `firm_id`    Unique firm-year identifier
25	         `ID`         Unique Blockholder-firm-year identifier
26	         `Shrsrc`     Source of blockholder data
27	         `shpct`      Percentage held by the blockholder, corrected for 
                          overlapping share holdings
28	         `sh_off`     The blockholder is an officer of the company
29	         `sh_dir`     Owned by a director of the company
30	         `ih_of_lb`   Block held indirectly 
                          (shares are attributed to other holder identified through `bh_ind`)
31	         `bh_ind`     Block held indirectly
32	         `bh_flag`    Counted as blockholder in adjusted data
33	         `aflin`      Block held by an affiliated entity
34	         `esopblk`    Held by an entity in its role as trustee for an employee stock
                          ownership plan
35	         `dir`        Owned by a non-officer director of the company
36	         `officer`    Owned by an officer (and possibly also a director) of the company
37	         `out`        Owner is neither an affiliated entity, nor an officer, nor a director,
                          nor an ESOP trustee
-----------  ------       -----------



\newpage


#Part 1: Data Prepartion

Take the following steps to create the data set that will be used for
further analysis:

\vspace{.2in}
1. Read in the data set appropriately. __(5)__

```{r}
FullData = read.csv("blockholders.csv", header = TRUE, sep = ",")
```

\vspace{.2in}
2. Remove all of the rows for which the variable `SH_name` is "Firm has no blockholder." __(3)__

```{r}
library(dplyr)
FilteredData = filter(FullData, FullData$SH_name!= "Firm has no blockholder")
# Insert R command(s) here.
```

\vspace{.2in}
3. The empty entries that appear in this data set are not actually "missing."
Instead, the columns with empty entries (e.g., `Partial`, `Full`, `PrefFlag`, etc.) are
_logical_ variables which indicate whether or not some condition holds.
For example, the variable `officer` indicates whether or not the blockholder is
an officer in the company.
In the original data file,
`1` indicates TRUE, and an empty entry indicates FALSE. Appropriately transform these
columns into R's logical format, i.e., filled with `TRUE` and `FALSE`. __(5)__

```{r}
FilteredData[is.na(FilteredData)] = 0
```

\vspace{.2in}
4. Appropriately format the two date columns into R's date format. __(3)__

```{r}
library("anytime")
FilteredData$mtgdate = anydate(as.character(FilteredData$mtgdate))
FilteredData$shrsrcd = anydate(as.character(FilteredData$shrsrcd))
```

\vspace{.2in}
5. Treat the variables `group2`, `Ticker` and `Shrsrc` as factors (categorical
variables), and `CompName`, `SH_name`, `firm_id`, and `ID` as character information. Make sure that R has them properly
represented, and convert as needed. __(4)__

```{r}
FilteredData[c("group2","Ticker", "Shrsrc")] = lapply(FilteredData[c("group2",
                                              "Ticker", "Shrsrc")] , factor)
FilteredData$CompName = as.character(FilteredData$CompName)
FilteredData$SH_name = as.character(FilteredData$SH_name)
FilteredData$firm_id = as.character(FilteredData$firm_id)
FilteredData$ID = as.character(FilteredData$ID)
#Should return True for below
is.factor(FilteredData$group2)
is.factor(FilteredData$Ticker)
is.factor(FilteredData$Shrsrc)

#Should return False for below
is.factor(FilteredData$CompName)
is.factor(FilteredData$SH_name)
is.factor(FilteredData$firm_id)
is.factor(FilteredData$ID)

```

\vspace{.3in}
_This is the end of Part One of the exam. Once you have finished Part One, create the .pdf file from this .Rmd file, and then submit both on Canvas. This will make the second part of the exam available to you._

_Remember that you cannot resubmit this first part._



